Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2016-12-16 ms_pre_post.Rmd
Edits up through Introduction made: left off now on last Model section. Switching to questions instead of hypotheses

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry are likely to be affected by direct negative selection on resulting phenotypes (i.e., costs) and hence on the net costs versus benefits stemming from inbreeding avoidance. Polyandrous females can pay costs of mate searching or mating, for example, because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice (e.g., reluctance to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that express post-copulatory choice might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs affecting pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly inbreeding strategy might become fixed in the population over generations, while alleles underlying the more costly inbreeding strategy might become extinct if their effects are both costly and made redundant by the less costly strategy.

Here we use individual-based modelling to set up *in silico* experiments, which test the hypothesis that, given strong inbreeding depression in offspring viability, evolution of post-copulatory inbreeding avoidance facilitates evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution such that only the less costly inbreeding avoidance strategy will be expressed in the long term. Finally, we test the hypothesis that evolution of one inbreeding avoidance strategy (i.e., pre-copulatory or post-copulatory) will be precluded if it is introduced into a population that is already fixed for the other inbreeding avoidance strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across different reproductive systems.

>>>>

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will depend not only on the magnitudes of positive indirect selection steming from reduced inbreeding depression in females' offspring, but also on the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). Polyandrous females can pay costs of mate searching or mating, for example, because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice and consequent reluctance to mate with relatives might pay costs of increased risk of complete mating failure or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that express post-copulatory choice might pay energetic costs associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly inbreeding strategy might become extinct if their effects are both costly and made redundant by the less costly strategy. However, such trajectories will also depend on the indirect benefits of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models are consequently required to understand and predict evolutionary outcomes.

<!--- This last sentence sounds good and strong, but it doesn’t actually follow logically from the rest of the paragraph – how do we know that we can’t predict or understand evolutionary outcomes with non-dynamic models, perhaps one that predicts long-term equilibria? This doesn’t seem obvious or well argued if we’re just talking about costs and benefits – what, specifically, can a dynamic model (I assume, a simulation, i.e., a model where things change over time) address that another type of model cannot? --->

We address three key questions for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given strong inbreeding depression. These questions include: (1) Does evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance facilitate evolution of polyandry that is otherwise severely constrained by increasing relatedness of additional mates? (2) How do costs associated with polyandry and pre-copulatory and post-copulatory inbreeding avoidance affect evolutionary outcomes and, in particular, the long-term persistence of reproductive strategies given cost assymetry? Finally, (3) how is selection for initially rare pre-copulatory or post-copulatory inbreeding avoidance affected if the other strategy of inbreeding avoidance is already fixed? By addressing these questions, we provide a framework for exploring the simultaneous evolution of multiple reproductive strategies through the use of individidual-based modelling, and we derive hypotheses amenable to empirical tests within and across taxa.
